Clicks and Cliques

### 2019-01-02 papersoul.Rnw
update references

This paper was part of 2013 Data Expo competition of the American Statistical Association. An overview of all papers and the data exposition can be
found in \cite{soul2}. A reproducible version of this paper is placed in the following  In GitHub repository: https://github.com/COSTDataExpo2013/DaSilvaAlvarez.

@@

\subsubsection{Sampling weights}
Statistical analysis of data from survey need to make a correct use of sampling weights. Depending of the source of information and the statistical goal, different sampling weights are used in several statistical methods or summaries.

As mentioned earlier, information from ACS corresponds to tables of community totals (e.g. Total of 18 years old women in each community). These variables were already expanded by Census Bureau, so there is no need to consider sampling weights again. For instance, proportions of age group by gender shown in Figure \ref{age} are computed directly from ACS data without explicit use of sampling weights.

When using data from Knight Foundation surveys, the sampling weights need to be taken into
account. There are two variables representing sampling weights that might be used considering if data from many communities are combined (PROJWT) or computations are done for each community separately (WEIGHT) \citep{full2010}. For instance, Figure \ref{pp1} includes data from all communities together so PROJWT is used as sampling weight, meanwhile Figure \ref{comu} shows proportions of attachment computed separately within each community  and therefore sampling weights correspond to WEIGHT variable. Additionally, when all three year survey data are used jointly, sampling weights are divided by three so the total aggregate of each variable still can be consider an estimate of the community total.

@@

The bottom panel of Figure \ref{eco} presents the U.S. GDP growth rate in 2008, 2009, and 2010. A similar pattern is observed in both panels, suggesting the economic dimension is less related with attachment during the economic crisis.
\begin{figure}[hbpt]
<<econ, dependson=c('data'), echo=FALSE,message=FALSE,warning=FALSE, fig.height=4.7>>=
svydesign(ids =~ 1, strata =~ QSB, weights =~ PROJWT, data = filter(soul,!is.na(PROJWT))) %>% 
  svyby( ~ ECONOMY + CCE, ~ YEAR, design = ., svyvar, keep.var = TRUE, na.rm = T)  %>% data.frame() %>% select(1:5) %>% 
  mutate(rho = V2/sqrt(V1*V4), gdp = c(-0.4,-3.5,3.0)) %>% 
  gather(variable, value,rho,gdp) %>% 
  select(YEAR, variable, value) %>% 
  mutate(variable =  factor(variable, labels = c('GDP Rate','Attachment-Economy correlation'))) %>% ggplot(aes(x = YEAR, y = value)) + 
  geom_line() + geom_point() +
  labs(x = "", y = "", 
       title = 'Evolution of Attachment-Economy correlation and GDP') +
  facet_grid(facets = variable~., scales = 'free') + 
  ggtitle('Evolution of Attachment-Economy correlation and GDP')

@@

There are two basic measures of variable importance in the \verb randomForest function within R, only the \emph{mean decrease in Gini} ($DG$) was used. Each time a variable is used within a random forest classifier is associated with a decrease in the impurity of the data nodes, which is a measure with the Gini impurity index. Important variables will achieve larger reduction of impurity, then $DG$ can be used as a variable importance measurement. Within each community explanatory variables were ranked according to $DG$. Rank 1 represents the variable associated with the largest reduction in the data impurity.

@@

\bibliographystyle{spr-chicago}

>>>>

This paper was part of 2013 Data Expo competition of the American Statistical Association. An overview of all papers and the data exposition can be
found in \cite{soul2}. A reproducible version of this paper is placed in the following GitHub repository: https://github.com/COSTDataExpo2013/DaSilvaAlvarez.

@@

\subsubsection{Sampling weights}
Statistical analyses of data from surveys need to make a correct use of sampling weights. Depending on the source of information and the statistical goal, different sampling weights are used in several statistical methods or summaries.

As mentioned earlier, information from ACS corresponds to tables of community totals (e.g. Total of 18 years old women in each community). These variables were already expanded by the Census Bureau, so there is no need to consider sampling weights again. For instance, proportions of age group by gender shown in Figure \ref{age} are computed directly from ACS data without explicit use of sampling weights.

When using data from Knight Foundation surveys, the sampling weights need to be taken into
account. There are two variables representing sampling weights that might be used considering if data from many communities are combined (PROJWT) or computations are done for each community separately (WEIGHT) \citep{full2010}. For instance, Figure \ref{pp1} includes data from all communities together so PROJWT is used as sampling weight, meanwhile Figure \ref{comu} shows proportions of attachment computed separately within each community  and therefore sampling weights correspond to WEIGHT variable. Additionally, when all three year survey data are used jointly, sampling weights are divided by three so the total aggregate of each variable still can be considered an estimate of the community total.

@@

The bottom panel of Figure \ref{eco} presents the U.S. GDP growth rate in 2008, 2009, and 2010. A similar pattern is observed in both panels, suggesting the economic dimension is less related with attachment during the economic crisis.
\begin{figure}[hbpt]
<<econ, dependson=c('data'), echo=FALSE,message=FALSE,warning=FALSE, fig.height=4.7>>=
svydesign(ids =~ 1, strata =~ QSB, weights =~ PROJWT, data = filter(soul,!is.na(PROJWT))) %>% 
  svyby( ~ ECONOMY + CCE, ~ YEAR, design = ., svyvar, keep.var = TRUE, na.rm = T)  %>% data.frame() %>% select(1:5) %>% 
  mutate(rho = V2/sqrt(V1*V4), gdp = c(-0.4,-3.5,3.0)) %>% 
  gather(variable, value,rho,gdp) %>% 
  select(YEAR, variable, value) %>% 
  mutate(variable =  factor(variable, labels = c('GDP Rate','Attachment-Economy correlation'))) %>% 
  ggplot(aes(x = YEAR, y = value)) + 
  geom_line() + geom_point() +
  labs(x = "", y = "", 
       title = 'Evolution of Attachment-Economy correlation and GDP') +
  facet_grid(facets = variable~., scales = 'free') + 
  ggtitle('Evolution of Attachment-Economy correlation and GDP') + 
  scale_x_continuous(breaks =  c(2008, 2009, 2010), minor_breaks = c(2008, 2009, 2010) )

@@

There are two basic measures of variable importance in the {\verb randomForest } function within R, only the \emph{mean decrease in Gini} ($DG$) was used. Each time a variable is used within a random forest classifier is associated with a decrease in the impurity of the data nodes, which is a measure with the Gini impurity index. Important variables will achieve larger reduction of impurity, then $DG$ can be used as a variable importance measurement. Within each community explanatory variables were ranked according to $DG$. Rank 1 represents the variable associated with the largest reduction in the data impurity.

@@

\bibliographystyle{bibstyle}
